Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Chad Bown, Nuno Limao, Giovanni Maggi, Anson Soderbery, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at SITE for very helpful comments and discussions.}

### 2017-01-11 draft.tex
simplify notation for e^R as well

\subsection{Perfect External Enforcement}
\label{sec:perfect}
I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement. Although it is widely agreed that this is unrealistic, this is an important baseline case. Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.

@@

In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_e^R$ for the case of endogenous political pressure with rigid bindings, as distinct from $\tau^R$ in the case of exogenous political pressure with rigid bindings. The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as $\tau^N(\ga(e)) \leq \tau_e^R$. Otherwise, it must set $\tau = \tau_e^R$ since the weak binding $\tau_e^R$ is externally enforced.

Knowing how the government will set the applied tariff, in the second stage the lobby makes its effort decision to maximize net profits according to Expression~\ref{eq:lv3}. If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_e^R$, the lobby will reduce its choice so that no effort is wasted. I label the lobbying effort choice that leads to the weak binding level as $e^R_{W,e}$ so that $\tau^N(\ga(e^R_{W,e})) = \tau_e^R$.

@@

The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding. The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$. The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}

@@

To understand how the optimal home tariff $\tau_e^R$ under the trade agreement is set, notice that it must maximize joint government welfare. As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
  \begin{equation}
		W_x \left(\ga(e),\tau) \right) + W_x^*\left(\tau \right)
	  \label{exp:1}
	\end{equation}
where the lobby will exert no more effort than that which produces the capped $\tau_e^R=\tau^N(\ga(e))$. Recall that the Nash tariff, $\tau^N(\ga(e))$ is the solution to the unilateral optimization problem the home government faces when choosing the applied tariff level in the final stage of the game. It is given by Expression~\ref{eq:nash}, which contains the mapping between the political pressure the lobby exerts and the government's unilateral tariff response. We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau_e^R$. This $e$ is that which is labeled $e^R_{W,e}$ above.

Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage. The governments get to choose the $\tau_e^R$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.

>>>>

\subsection{Perfect External Enforcement}
\label{sec:perfect}
I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement. Although this is an unrealistic assumption, it establishes an important baseline case. Section~\ref{sec:self} adds the requirement that trade agreements be self-enforcing.

@@

In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_e^R$ for the case of endogenous political pressure with rigid bindings, as distinct from $\tau^R$ in the case of exogenous political pressure with rigid bindings. The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as $\tau^N(\ga(e)) \leq \tau_e^R$. Otherwise, the government must set $\tau = \tau_e^R$ since the weak binding $\tau_e^R$ is externally enforced.

Knowing how the government will set the applied tariff, in the second stage the lobby makes its effort decision to maximize net profits according to Expression~\ref{eq:lv3}. If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_e^R$, the lobby will reduce its choice so that no effort is wasted. I label the lobbying effort choice that leads to the weak binding level as $e^R$ so that $\tau^N(\ga(e^R)) = \tau_e^R$.

@@

The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding. The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$. The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}

@@

To understand how the optimal home tariff $\tau_e^R$ under the trade agreement is set, notice that it must maximize joint government welfare. As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
  \begin{equation}
		W_x \left(\ga(e),\tau) \right) + W_x^*\left(\tau \right)
	  \label{exp:1}
	\end{equation}
where the lobby will exert no more effort than that which produces the capped $\tau_e^R=\tau^N(\ga(e))$. Recall that the Nash tariff, $\tau^N(\ga(e))$ is the solution to the unilateral optimization problem the home government faces when choosing the applied tariff level in the final stage of the game. It is given by Expression~\ref{eq:nash}, which contains the mapping between the political pressure the lobby exerts and the government's unilateral tariff response. We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau_e^R$. This $e$ is that which is labeled $e^R$ above.

Ex-post lobbying will constrain the governments in that the $\ga\left(e^R\right)$ they experience will be determined through the unilateral maximization problem in the last stage. The governments get to choose the $\tau_e^R$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.
